---
title: "What We Talk About When We Talk About Data Science on Medium.com"
output: 
  flexdashboard::flex_dashboard:
    source_code: embed
    theme: journal
runtime: shiny

---

```{r setup, include=FALSE}
## Import libraries
library(flexdashboard)
library(shinyWidgets)
library(igraph)
library(ggraph)

## Set seed for ggraph package
set.seed(2016)
```


 

Explore top article title keywords
=========================================


Column {.sidebar}
-----------------------------------------------------------------------

 
```{r}
sliderInput("year", HTML("<b>Select the year of publication:</b>"), 
            min = 2013, max = 2019, value = 2013,
            sep="", animate=animationOptions(interval = 4000, loop = FALSE, playButton = NULL,
  pauseButton = NULL))
```

<br><hr><br>


```{r}
knobInput(
  "knob",
  HTML("<b>Select the top X most frequently appearing bigrams:</b>"),
  40,
  min = 30, max = 70,
  step = 5, angleOffset = 0, angleArc = 360,
  cursor = TRUE,
  thickness = NULL,
  lineCap = "round",
  displayInput = TRUE,
  displayPrevious = FALSE,
  rotation = "clockwise",
  fgColor = NULL,
  inputColor = NULL,
  bgColor = NULL,
  readOnly = FALSE,
  skin = NULL,
  width = NULL,
  height = NULL,
  immediate = FALSE
)
```

Column {data-width=350 .tabset .tabset-fade}
-----------------------------------------------------------------------

### Introduction

In this dashboard, you can explore the most frequently appearing bigrams in the titles of Medium.com articles tagged with "Data science" published between 2013 and late November of 2019.

Please see [here](http://rpubs.com/nchelaru/medium_ds_nlp) for a detailed run-down of the workflow for web scraping and data preparation.


### General workflow

Here, I used the [workflow](https://www.tidytextmining.com/ngrams.html#counting-and-correlating-pairs-of-words-with-the-widyr-package) presented in the excellent reference *[Text Mining with R](https://www.tidytextmining.com/)* for visualizing bigram relationships for a body of text.

```r
## Import libraries
library(igraph)
library(ggraph)

## Set seed for ggraph package
set.seed(2016)

## Create graph
bigram_graph <- bigram_counts %>% 
                    graph_from_data_frame()

a <- grid::arrow(type = "closed")

ggraph(bigram_graph, layout = "fr") + 
    geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, 
                   arrow = a, end_cap = circle(0.07, "inches")) + 
    geom_node_point(color = "lightblue", size = 5) + 
    geom_node_text(aes(label = name), vjust = 1, 
                   hjust = 1, repel = TRUE) + 
    theme_void()
```

Column {data-width=650}
-----------------------------------------------------------------------

### Visualize network of top occurring bigrams
 
```{r}
renderPlot({
  link <- sprintf("https://github.com/nchelaru/scrapy/raw/master/data/processed/%s_title_bigram_counts.csv", input$year)

  ## Import data
  bigram_counts <- read.csv(link)
  
  bigram_counts <- head(bigram_counts[order(-bigram_counts$n),], input$knob)
  
  ## Create graph
  bigram_graph <- bigram_counts %>%
    graph_from_data_frame()
  
  a <- grid::arrow(type = "closed")
          
  ggraph(bigram_graph, layout = "fr") +
    geom_edge_link(aes(edge_alpha = n), show.legend = FALSE,
                   arrow = a, end_cap = circle(.05, 'inches')) +
    geom_node_point(color = "lightblue", size = 5) +
    geom_node_text(aes(label = name), vjust = 1, hjust = 1, repel = TRUE, size=8) +
    theme_void() 
})

```


Session info
=========================================
```{r}
sessionInfo()
```

